\Title

### 2021-05-31 latex/paper/main.tex
Changed some sentence structure in the proof of the Theorem

Suppose for a moment that \(I = 0\) in \eqref{eqn:acrobot-constraint}, i.e.,
that the legs stay fully extended.
The acrobot becomes a nominal pendulum with two masses, whose total mechanical
energy is
\begin{equation}\label{eqn:acrobot-energy}
    E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
    .
\end{equation}
The upright equilibrium of this pendulum is located at \((q_u,p_u) = (\pi,0)\).
Imagine the pendulum hits the bottom of the swing arc with momentum 
\(p_u \neq 0\). 
To reach the upright equilibrium, this momentum must be
\(p_u = \pm\sqrt{60m^2gl^3}\) because \(E(\pi,0) = E(0,\pm\sqrt{60m^2gl^3})\).
If the momentum is smaller in magnitude, the acrobot will oscillate; 
if it is larger, the pendulum will rotate around the bar.

@@

Recall that we defined the sets \(\mathcal{O}_1\) and
\(\mathcal{O}_2(\bar{\rho})\) used in Theorem \ref{thm:acrobot-energy-stabilization}
via the mechanical energy \eqref{eqn:acrobot-energy} of the nominal
pendulum associated with a simplified acrobot.
We obtained this nominal pendulum by setting \(I = 0\) in
\eqref{eqn:acrobot-constraint}.
Recall that the boundary between oscillations and rotations of the nominal
pendulum is obtained by finding the level set \(E_\pi\) of the nominal energy
corresponding to the upright equilibrium \((q_u,p_u) = (\pi,0)\).

@@

%/========== Proof ==========/%
\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
The sets \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\) are sublevel sets of
the nominal pendulum's energy \(E(q_u,p_u)\) described in
\eqref{eqn:acrobot-energy}.
To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will find conditions
under which the acrobot gains energy on an open ``energy annulus".
That is, given two energy levels \(E_0 \geq 0\) and \(E_1 > E_0\), we define the
energy annulus to be the domain
\begin{equation*}
    \mathcal{D} := \{(q_u,p_u) \in \SxR \mid \\
    E_0 < E(q_u,p_u) < E_1\}
    .
\end{equation*}
The only condition we place on \(E_0\) and \(E_1\) is that they both correspond
to oscillations or both to rotations; that is, if \(E_0 < E(\pi,0)\), then 
\(E_1 \leq E(\pi,0)\).

Here is the fundamental idea of the proof: for certain open subsets
\(D \subset \mathcal{D}\), one can find a change of coordinates into 
pseudo-polar coordinates adapted to level sets of \(E\).
In other words, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
a diffeomorphism \(T(q_u,p_u) = (r,\theta)\), where \(r\)
chooses a level set of \(E\) and \(\theta\) chooses a point on that level set.

We will describe the explicit forms of these transformations later. 
Importantly, we will show that the dynamics of \((r,\theta)\) satisfy 
\(\dot{r} = 0\) and \(\dot{\theta} > 0\) for the nominal pendulum (i.e., when
\(I = 0\)).
By continuity of the constrained dynamics with respect to \(I\), we can find a
small enough \(I > 0\) so that \(\dot{\theta} > 0\) for the acrobot.
This property will enable us to perform a time-scaling transformation 
\(t = t(\theta)\), and study how the pseudo-radius \(r\) evolves as a function
of the pseudo-angle \(\theta\), rather than as a function of time.
If \(r(\theta)\) increases on average, then \(E\) must also be increasing
on average and the constrained dynamics must be gaining energy on \(D\).

We formalize this idea in the following lemma.
Then we will find the relevant subset \(D \subset \mathcal{D}\)
and the coordinate transformations \((q_u,p_u) \mapsto (r,\theta)\) which
satisfy the conditions of the lemma, thereby proving the acrobot gains energy on
each set.

\begin{lemma}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    parameterized by \((q_u,p_u) \in \SxR\).
    Let \(D \subset \SxR\) be open and 
    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    satisfying \(\dot{r} \equiv 0\) and \(\dot{\theta} > 0\) when
    \(I = 0\).
    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
    \((\partial g/\partial I)(\theta,r,0) \equiv 0\).
    Define \(b(\theta,r_0) = (\partial g/\partial r)(\theta,r_0,0)\);
    assume it is \(2\pi\)-periodic in \(\theta\) for fixed \(r_0\),
    and that there exists \(\epsilon > 0\) such that 
    \[
        \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
        ,
    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
    for all \(I \in [-I^\star, 0[\).
\end{lemma}
\begin{proof}
    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, ]0,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE 
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \frac{\dot{r}}{\dot{\theta}} =
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    where the prime denotes differentiation by \(\theta\) and 
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

@@

    Proving the constrained dynamics gain energy on \(D\) is equivalent to
    proving \(\hat{r}_{n}\) eventually reaches \(\overline{r}\).
    Thanks to a perturbation analysis from \cite[Theorem 10.1]{khalil_nonlinear}, 
    we know there exists some \(I_2 \in ]0,I_1]\) and \(\delta > 0\)
    such that \(\left|R(2\pi,r_n,I)\right| \leq \delta I^2\) for all 
    \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0)\) is uniformly
    bounded below by \(\epsilon\), we can find a lower bound on
    \eqref{eqn:r-discrete}:
    \[
        r_{n+1} \geq r_n + I\epsilon - \delta I_2^2
        .
    \]
    Choosing \(I^\star \in ]0,I_2]\) small enough makes 
    \(I\epsilon - \delta I_2^2 \geq \gamma > 0\) for all \(I \in ]0,I^\star]\).
    Therefore, \(r_{n+1} \geq r_n + \gamma \geq r_0 + n\gamma\), which means all
    solutions of \eqref{eqn:r-discrete} flow towards \(\overline{r}\).
    We conclude that, for \(I \in ]0, I^\star]\), the constrained dynamics gain
    energy on \(D\).
    Changing the sign of \(I\) means \(r_{n+1} \leq r_n - n\gamma\), in which
    case the constrained dynamics lose energy on \(D\).
\end{proof}
\textbf{TODO: Prove origin is unstable, then apply lemma to O1 and O2 and show
that orbits go through Epi}
%/========== Conclusion ==========/%
\section{Conclusion}\label{sec:conclusion}

>>>>

Suppose for a moment that \(I = 0\) in \eqref{eqn:acrobot-constraint}, i.e.,
that the legs stay fully extended.
The acrobot becomes a nominal pendulum with two masses, whose total mechanical
energy is
\begin{equation}\label{eqn:nominal-energy}
    E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
    .
\end{equation}
The upright equilibrium of this pendulum is located at \((q_u,p_u) = (\pi,0)\).
Imagine the pendulum hits the bottom of the swing arc with momentum 
\(p_u \neq 0\). 
To reach the upright equilibrium, this momentum must be
\(p_u = \pm\sqrt{60m^2gl^3}\) because \(E(\pi,0) = E(0,\pm\sqrt{60m^2gl^3})\).
If the momentum is smaller in magnitude, the acrobot will oscillate; 
if it is larger, the pendulum will rotate around the bar.

@@

Recall that we defined the sets \(\mathcal{O}_1\) and
\(\mathcal{O}_2(\bar{\rho})\) used in Theorem \ref{thm:acrobot-energy-stabilization}
via the mechanical energy \eqref{eqn:nominal-energy} of the nominal
pendulum associated with a simplified acrobot.
We obtained this nominal pendulum by setting \(I = 0\) in
\eqref{eqn:acrobot-constraint}.
Recall that the boundary between oscillations and rotations of the nominal
pendulum is obtained by finding the level set \(E_\pi\) of the nominal energy
corresponding to the upright equilibrium \((q_u,p_u) = (\pi,0)\).

@@

%/========== Proof ==========/%
\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:proof}
Theorem \ref{thm:acrobot-energy-stabilization} proves that the acrobot gains
energy on \(\mathcal{O}_1\) and \(\mathcal{O}_2(\bar{\rho})\), which are
sublevel sets of the nominal pendulum's energy \eqref{eqn:nominal-energy}.
%To prove Theorem \ref{thm:acrobot-energy-stabilization}, we will find conditions
%under which the acrobot gains energy on an open ``energy annulus",
%\begin{equation*}
%    \mathcal{D} := \{(q_u,p_u) \in \SxR \mid \\
%    E_0 < E(q_u,p_u) < E_1\}
%    ,
%\end{equation*}
%where \(E_1 > E_0\) are energy values, which we will choose later.

Here is the fundamental idea of this proof: for certain subsets
\(D \subset \SxR\), one can find a change of coordinates into 
pseudo-polar coordinates adapted to level sets of \(E\) 
(see Mohammadi \etal \cite{dynamic_vhcs_stabilize_closed_orbits}).
In other words, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, one can find
a diffeomorphism \(T(q_u,p_u) = (r,\theta)\), where \(r\)
chooses a level set of \(E\) and \(\theta\) chooses a point on that level set.

We will explicitly calculate these transformations later;
what matters is that the dynamics of \((r,\theta)\) will satisfy 
\(\dot{r} = 0\) and \(\dot{\theta} > 0\) when \(I = 0\).
By continuity of the constrained dynamics with respect to \(I\), this implies 
there exists a small enough value \(I > 0\) so that \(\dot{\theta} > 0\) for the
acrobot.
This property will enable us to perform a time-scaling transformation 
\(t = t(\theta)\), and study how the pseudo-radius \(r\) evolves as a function
of the pseudo-angle \(\theta\), rather than as a function of time.
If \(r(\theta)\) increases on average, then \(E\) must also be increasing
on average and the constrained dynamics must be gaining energy on \(D\).
This idea is formalized in the following lemma.

\begin{lemma}\label{lemma:energy-regulation-D}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    parameterized by \((q_u,p_u) \in \SxR\).
    Let \(D \subset \SxR\) be open and 
    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    satisfying \(\dot{r} \equiv 0\) and \(\dot{\theta} > 0\) when
    \(I = 0\).
    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
    \((\partial g/\partial I)(\theta,r,0) \equiv 0\).
    Define \(b(\theta,r_0) = (\partial g/\partial r)(\theta,r_0,0)\);
    assume it is \(2\pi\)-periodic in \(\theta\) for fixed \(r_0\),
    and that there exists \(\epsilon > 0\) such that 
    \[
        \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
        ,
    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
    for all \(I \in [-I^\star, 0[\).
\end{lemma}
\begin{proof}
    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, ]0,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE 
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \frac{\dot{r}}{\dot{\theta}} =
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    where the prime denotes differentiation by \(\theta\) and 
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

@@

    Proving the constrained dynamics gain energy on \(D\) is equivalent to
    proving \(\hat{r}_{n}\) eventually reaches \(\overline{r}\).
    Thanks to a perturbation analysis from \cite[Theorem 10.1]{khalil_nonlinear}, 
    we know there exists some \(I_2 \in ]0,I_1]\) and \(\delta > 0\)
    such that \(\left|R(2\pi,r_n,I)\right| \leq \delta I^2\) for all 
    \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0)\) is uniformly
    bounded below by \(\epsilon\), we can find a lower bound on
    \eqref{eqn:r-discrete}:
    \[
        r_{n+1} \geq r_n + I\epsilon - \delta I_2^2
        .
    \]
    Choosing \(I^\star \in ]0,I_2]\) small enough makes 
    \(I\epsilon - \delta I_2^2 \geq \gamma > 0\) for all \(I \in ]0,I^\star]\).
    Therefore, \(r_{n+1} \geq r_n + \gamma \geq r_0 + n\gamma\), which means all
    solutions of \eqref{eqn:r-discrete} flow towards \(\overline{r}\).
    We conclude that, for \(I \in ]0, I^\star]\), the constrained dynamics gain
    energy on \(D\).
    Changing the sign of \(I\) means \(r_{n+1} \leq r_n - n\gamma\), in which
    case the constrained dynamics lose energy on \(D\).
\end{proof}

\textbf{TODO: Prove origin is unstable, then apply lemma to O1 and O2 and show
that orbits go through Epi}
%/========== Conclusion ==========/%
\section{Conclusion}\label{sec:conclusion}
